---
title: "Lab 1.2 - The Seattle Pets Dataset"
author: "Trey Tipton"
date: "Fall 2022"
output: html_document
---

This RMarkdown document presents a preliminary analysis of Seattle&rsquo;s [pet licenses dataset](https://data.seattle.gov/Community/Seattle-Pet-Licenses/jguv-t9rb). It interleaves text, code, and code output. 

## Loading the Dataset

The following code *chunk* loads the tidyverse library, reads the CSV file containing the Seattle pet license dataset, and saves that dataset under the name `seattle_pets`. The code assumes that the CSV file is loaded in a sub-directory, `data`, of the directory that contains this document.

```{r setup, message=FALSE}
library(tidyverse)
seattle_pets <- read_csv("data/Seattle_Pet_Licenses.csv")
```

## Viewing the Structure of the Dataset

We can now view the dataset as it is stored in R.

```{r pets}
seattle_pets
```

Based on the information provided in the previous section, we can see that the pets dataset contains how many of the following:

- Pets (i.e., rows, a.k.a. *records*):  44,565
- Variables (i.e., columns, a.k.a. *fields*):  7

## Analyzing the Dataset

We can now count the number of each species using the `count()` function.

```{r count-by-species}
count(seattle_pets, Species, sort=TRUE)
```

The last chunk of code tells us the number of records in the data set that are each species, of which there are four.. It shows that the species that shows up the most is dogs, then cats. then goats, and then pigs.

```{r count-by-name}
count(seattle_pets, `Animal's Name`, sort=TRUE)
```

This code chunk tells us the most popular animal names, the most common of which being Luna.


