---
title: "Stat 245 -- 3+ Variable Plots"
author: "Trey Tipton"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 4.2
    fig_width: 6
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

```{r}
searches <- read_csv('https://sldr.netlify.app/data/election_searches.csv')
```

I chose one quantitative variables, vote sway, and three categorical, their employment status, whether they trusted the search results, and sex.

```{r}
gf_boxplot(Vote_Sway ~ Trust_Search|Employment_Status, data = searches, color = ~Sex, ylab = "Vote Sway")
```

A pattern I see in these boxplots is that every single median (for male and female) is at zero. This does not tell us that there is no vote sway on average, but that the vote sway on each side is even. Given the fact that there is outliers on both sides of each boxplot, I suspect that there was actually quite a bit of vote sway, they just happened to even out. Sex does not seem to have much of an effect on vote sway, however, "other" has the only boxplots that are not at a median of zero, although this is likely because of a small smaple size of the sex "other" in this dataset. Employment and whether they trusted the results also does not seem to have much of an effect on vote sway. The one slightly different one is employed males that suspected bias in the search results; their range is a lot larger, but they also have less outliers and this is likely why.


  