---
title: "Stat 344 Final"
author: "Trey Tipton"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too
require(circular)
require(maxLik)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### 0. Academic Honesty Statement

I, Trey Tipton, affirm on my honor that this submission contains my own work and I did not get help from, or give help to, another person.

### 1. Circular Oxen

```{r}
ox <- read.csv('https://sldr.netlify.app/data/ox.csv') |>
  mutate(state = factor(state)) |>
  na.omit()

gf_path(ys ~ xs, data = ox, color = 'grey44') |>
  gf_point(color = ~state) |>
  gf_labs(x = 'X position', y = 'Y position')
```






## A. von Mises MLE

```{r, warning = FALSE}
ll_vm <- function(theta, x){
  mu <- theta[1]
  kappa <- theta[2]
  if (kappa < 0) return(NA)
  dvonmises(x, mu = mu, kappa = kappa, log = TRUE)
}

maxLik(ll_vm, start = c(mu = 1, kappa = .1), x = ox$angle)
```

Using maximum-likelihood estimation to fit a von Mises distribution to the angle data, the fitted parameter estimates are as follows:

$\mu = 0.349443 , \kappa = 0.1005169$.

```{r, warning = FALSE}
gf_dhistogram(~angle, data = ox) %>%
  gf_labs(x = 'Angle (radians)', y = 'Density') %>%
  gf_dist('vonmises', mu = 0.349443, kappa = 0.1005169)
```

## B. von Mises - uniform

```{r, warning = FALSE}
maxLik(ll_vm, start = c(mu = -1, kappa = 0), fixed = 2, x = ox$angle)

ll_unif <- function(theta, x){
  mu <- theta[1]
  dvonmises(x, mu = mu, kappa = 0, log = TRUE)
}

maxLik(ll_unif, start = c(mu = 10), x = ox$angle)

mu_starting_val <- c(-1, 0, 1, 2, 3, 4, 5, 6)
ll_estimation <- c(-1, 0, 1, 2, 3, 4, 5, 6)

gf_point(mu_starting_val ~ ll_estimation)%>%
  gf_labs(x = "Log-Likelihood Estimation", y = "mu Starting Value")
```

It appears that when $\kappa$ is zero, the log-likelihood estimation for $\mu$ is just the starting value for $\mu$. This is likely because when $\kappa$ is zero, the distribution is uniform (perfectly circular), meaning that the mean ($\mu$) is fixed based on what starting value you use, because the maxLik function cannot optimize for $\mu$ when the data is circular.


## C. von Mises test 

Hypotheses:

$H_0$: $\Omega$ = $\Omega_0$, $\kappa$ is equal to zero and the distribution of the oxen movement angles is therefore uniform.

$H_a$: $\Omega$ = $\Omega_1$, $\kappa$ is greater than zero and the oxen movement angles fit better to some sort of von Mises distribution.

where $\theta$ = <$\mu$, $\kappa$>,

$\Omega_0$ = {$\theta$ | $\kappa$ = 0}, and

$\Omega_1$ = {$\theta$ | $\kappa$ > 0}}

Test Stat:

Using the maximum likelihood estimations from the previous question, we got two values, one from the von Mises MLE and one from the uniform MLE (von Mises but with $\kappa$ fixed at zero):

von Mises Log-Likelihood: -2589.687 

Uniform Log-Likelihood: -2593.245

We use those for our W test statistic:

$W = 2*(l(\Omega_0) - l(\Omega_1))$

```{r}
w <- 2*(-2589.687  - -2593.245 )
1 - pchisq(w, 1)
```

With a low p-value of 0.00764, we reject the null hypothesis that a uniform distribution generates the data and that $\kappa$ is zero. Therefore, it is likely that $\kappa$ is greater than zero and a von Mises distribution fits the data better, since we do not have enough evidence to say that $\kappa = 0$.

## D. Goodness!


Hypotheses:

$H_0: \Omega_0 =${$\theta | \kappa > 0$}, where $\theta$ = <$\mu$, $\kappa$>, A von Mises distribution generates the angle data for the ox data set.


$H_a:$ A von Mises distribution does not generate the angle data for the ox data set.

Test Statistic:

$G = -2 * log(\lambda)$ = $2 * \sum o_i * log(\frac{o_i}{e_i})$


```{r, warning = FALSE}
min(ox$angle)
max(ox$angle)
angle_breaks <- c(-pi+.000001, -(3*pi)/4, -pi/2, -pi/4, 0, pi/4, pi/2, (3*pi)/4, pi)
ox <- ox %>%
  mutate(binned_angles = cut(angle, breaks = angle_breaks))

tally(~binned_angles, data = ox)

count_dat <- data.frame(tally(~binned_angles, data = ox))
count_dat


count_dat <- count_dat %>%
  mutate(probs = diff(pvonmises(angle_breaks, mu = 0.349443, kappa = 0.1005169, 
                                from = pi+.000000001)), e = sum(count_dat$Freq) * probs)
count_dat$e


G_angle <- 2 * sum( count_dat$Freq * log( count_dat$Freq / count_dat$e))
G_angle

1 - pchisq(G_angle, df = nrow(count_dat) - 1 - 2)

pearson_angle <- sum(((count_dat$Freq - count_dat$e)^2) / count_dat$e)

1 - pchisq(pearson_angle, df = 2)
```

Both of the p-values we computed from the Likelihood Ratio Test Statistic and the Pearson chi-square test stat are very low at 5.942794e-06 and 1.831628e-07, therefore we reject the null hypothesis that a von Mises distribution is a good fit to the data. This means that although it is a better fit than a uniform distribution, we do not have enough evidence to say that the angle data are formed by a von Mises distribution.



### 2. More Oxen

## A. Regression - Fit 

```{r, warning = FALSE}
ox_lm <- glm(state ~ altitude + time, data = ox, family = binomial(link = 'logit'))
msummary(ox_lm)
```

Model Equation:

$$logit(p_{meandering}) = log \bigg{(} \frac{p_{meandering}}{(1-p_{meandering})} \bigg{)} = 1.36518 + 0.00825x_{altitude} + 0.82901x_{time}$$

$$y_{state} \sim {\sf Binom}(1, p_{meandering})$$

where $p_{meandering}$ is the probability that an ox is meandering, and $x_{time}$ = 2 if it is nighttime and 1 if daytime.



## B. Ox at night 

```{r}
p <- 1/(1 + exp(-(1.36518 + 0.82901*2)))
p
```


The probability that an ox would be meandering when at sea level ($x_{altitude}$ = 0) and at midnight ($x_{timenight}$ = 2) is 0.9536113.


## C. Ox at night again 

```{r}
odds_night <- 0.9536113/(1 - 0.9536113)
odds_night
p_day <- 1/(1 + exp(-(1.36518 + 0.82901)))
odds_day <- p_day/(1-p_day)
p_day
odds_day

st.err <- sqrt(diag(vcov(ox_lm)))
st.err

odds <- -(odds_night - odds_day)

odds

odds + c(-1, 1)*st.err[3]*qnorm(0.975)
```

The odds of meandering for the sea-level ox decreases by about 11.58424 as time advances from midnight to daytime. From this we can create a confidence interval: we are 95% confidence that the change in odds of meandering at sea-level from midnight to daytime is contained in the interval (-12.10306, -11.06543).



### 3.Yet more oxen

## A. Fit a model 

```{r}
ox <- ox %>%
  mutate(time2 = ifelse(time=="day",1,0))


ll_mod <- function(theta, x){
  beta0 <- theta[1]
  beta1 <- theta[2]
  beta2 <- theta[3]
  beta3 <- theta[4]
  beta4 <- theta[5]
  sigma <- theta[6]
  resids <- ox$step - (beta0 + beta1*ox$time2 + beta2*ox$temp + beta3*ox$altitude + beta4*ox$xs)
  if (sigma < 0) return(NA)
  dnorm(resids, mean = 0, sd = sigma, log = TRUE)
}


maxLik(logLik = ll_mod, start = c(beta0 = 1, beta1 = 1, beta2 = -1, beta3 = 0, beta4 = -1, sigma = 1), x = ox)
```

$$y_{step} = 144.2272 + 64.61116x_{time} - 6.31048x_{temp} -0.7636456x_{altitude} + 41.22386x_{xs} + \epsilon$$

$$\epsilon \sim N(0, 0.9229477)$$
where $y_{step}$ is ox's step length and $x_{time}$ is zero if it is night, and 1 if it is day.

## B. Select

```{r}
ox_model <- lm(step ~ time2 + temp + altitude + xs, data = ox)
msummary(ox_model)
```

Backwards Step-wise Selection: we will do an Anova test on the full model to see which predictor produces the largest large p-value, and continue doing so until all p-values are small.

```{r}
car::Anova(ox_model)
```

From the full model, there is no large p-values, so each predictor plays a significant role in the model.  The best model includes time, temp, altitude, and xs as predictors for step length.


## C. Predict

```{r}
new_data <- data.frame(altitude = 0, time2 = 0, xs = .5, temp = -10)
conf_int <- predict(ox_model, newdata = new_data, 
        interval = 'prediction',
        level = 0.95)
conf_int
```

A 95% prediction interval interval for the distance (meters) moved in the next hour for an ox at sea level, at midnight, when it was -10 degrees, and at an xs position of 0.5 is (-402.2231  , 783.376). Since we know that the ox can only move a positive distance, we can predict with 95% confidence that the ox will move between 0 and 783.376 meters in the next hour.


