---
title: "Stat 344 -- HW 6"
author: "Trey Tipton"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### Problem 5.46 d.)

```{r}
classes <- c(9, 39, 22, 35, 28)

lambda <- mean(classes)

lambda

```
We want a gamma prior with a mean of 26.6


```{r}
gf_dist(dist = 'gamma', 
        params = list(shape = 21.28,
                      rate = .8)) |>
  gf_labs(title = 'Prior')
```

```{r}
alpha <- 21.28
theta <- .8 

sum(classes) + alpha

length(classes) + theta
```


$$g(\lambda) = \text{Gamma}(\alpha = 154.28, \theta = 5.8)$$

```{r}
gf_dist(dist = 'gamma', 
        params = list(shape = 154.28,
                      rate = 5.8)) |>
  gf_labs(title = 'Posterior')
```


```{r}
set.seed(032601)
post_sample <- rgamma(n = 10000, shape = 154.28, rate = 5.8)
```




```{r}
cdata(~post_sample, p = 0.95)
```

A 95% credible interval for lambda for our data is (22.54, 30.98). This means that we are 95% confident that this interval contains the true value of lambda (also known as the mean) for a poisson distribution that our data came from.


### Problem 5.47
a.)
```{r, results = "hide"}
set.seed(032601)
Positions <- rgeo(25)
leaflet_map(position = Positions, mark = TRUE)
```
Observed: land = 5, water = 20.


Using uniform prior of Unif(0,1):

prior: $g(\pi) = 1$
likelihood: $f(x|\pi) = \pi^a(1-\pi)^b$

which means our posterior is: $h(/pi|x) = (1)*(\pi^a(1-\pi)^b)$, so $\pi$ ~ $Beta(5 + 1, 20 + 1)$

```{r}
set.seed(032601)
post_unif <- rbeta(n = 10000, shape1 = 6, shape2 = 21)
cdata(~post_unif, p = 0.90)
```
A 90% credible interval for the proportion using a uniform prior is (0.106, 0.361).

```{r}
gf_dist(dist = 'beta', 
        params = list(shape1 = 4,
                      shape2 = 8)) |>
  gf_labs(title = 'Posterior')
```

Using Beta prior of Beta(4,8):

prior: $g(\pi) = \pi^{\alpha - 1}(1 - \pi)^{\beta - 1}$
likelihood: $f(x|\pi) = \pi^a(1-\pi)^b$

which means our posterior is: $h(\pi|x) = \pi^{\alpha-1+a}(1-\pi)^{\beta-1+b}$, so $\pi$ ~ $Beta(\alpha + a, \beta + b)$

$\pi$ ~ $Beta(9, 28)$

```{r}
set.seed(032601)
post_beta <- rbeta(n = 10000, shape1 = 9, shape2 = 28)
cdata(~post_beta, p = 0.90)
```
A 90% credible interval for the proportion using a beta prior is (0.138, 0.364).


b.)

```{r, results = "hide"}
set.seed(032641)
Positions <- rgeo(25)
leaflet_map(position = Positions, mark = TRUE)
```
Observed land=8, water=17.

Using Beta prior of Beta(9,28) (from last posterior):

Our new prior is $g(\pi)$ = $h(\pi|x) = \pi^{\alpha-1+a}(1-\pi)^{\beta-1+b}$.
likelihood: $f(x|\pi) = \pi^a(1-\pi)^b$

New posterior is: $h(\pi|x) = \pi^{\alpha-1+2a}(1-\pi)^{\beta-1+2b}$.

so $\pi$ ~ $Beta(\alpha + 2a, \beta + 2b)$

$\pi$ ~ $Beta(9 + 2(8),28 + 2(17))$

```{r}
set.seed(032601)
post_beta2 <- rbeta(n = 10000, shape1 = 25, shape2 = 62)
cdata(~post_beta2, p = 0.90)
```

Final using original posterior with 50 observations:

Using Beta prior of Beta(9,28):

posterior is the same: $h(\pi|x) = \pi^{\alpha-1+2a}(1-\pi)^{\beta-1+2b}$.

so $\pi$ ~ $Beta(\alpha + 2a, \beta + 2b)$

$\pi$ ~ $Beta(9 + 2(13),28 + 2(37))$

```{r}
set.seed(032641)
post_beta2 <- rbeta(n = 10000, shape1 = 35, shape2 = 102)
cdata(~post_beta2, p = 0.90)
```

c.)
```{r}
grid <- expand.grid(pi = seq(from = 0, by = 0.001, to = 1))
```

```{r}
a <- 13
b <- 37
alpha <- 9
beta <- 28

grid <- grid |>
  mutate(prior = pi^(alpha - 1 + 2*a)*(1-pi)^(beta - 1 + 2*b), likelihood = dbinom(x = a, size = a + b, prob = pi), posterior = prior * likelihood)

numerical_posterior <- with(grid, sample(x = pi, size = 1e6, prob = posterior, replace = TRUE))

gf_dhistogram(~numerical_posterior) %>%
  gf_dist('beta', shape1 = 34, shape2 = 101)
```

The Numerically obtained posterior sample distribution of the proportion seems to agree with the analytically obtained posterior distribution.






