---
title: "Stat 344 -- HW 3"
author: "Trey Tipton"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### Problem 5.24
a.) The likelihood ratio test statistic is $\lambda$ = $L(0)$/$L(\hat{\theta})$
see paper work

b and c.)
```{r}
x <- c(.64, .92, .73, .96, .98, .33, .8, .96, .81, .76, 
       .98, .75, .87, .82, .44, .96, .61, .32, .67, .98, 
       .96, .88, .85, 1.0, .86, .88, .8, .83, .64, .5)

n <- length(x)
theta_hat <- (-n/(sum(log(x)))) -1
theta_hat

lambda <- 1/(prod((theta_hat +1)*(x^(theta_hat))))
lambda
w_test <- -2*(log(lambda))
w_test
## w should equal w_test 
w <- 2*(30*log(theta_hat + 1) + theta_hat*sum(log(x)))
w
1 - pchisq(w, 1)
```
d.)
```{r}
W <- function(x, theta_0 = 0){ 
n <- length(x)
theta_hat <- (-n/(sum(log(x))) - 1)
2*(30*log(theta_hat + 1) + theta_hat*sum(log(x)))
}

rand_data <- do(1000)*runif(30, 0, 1)

statTally(sample = x, rdata = rand_data, FUN = W)
```
Our p-value for 1000 replications is 7.23e-12

e.) Both of the p-values are very small, leading us to the conclusion that we should reject the nully hypothesis that $\hat{\theta}$=0.


### Problem 5.30

a.) on paper, $\hat{\theta}_{MLE}$=0.325

b.)
```{r}
w <- -2 * ((1997*log(.25*(2+.05)) + 906*log(.25*(1-.05)) + 904*log(.25*(1-.05)) + 32*log(.25*(.05))) - (1997*log(.25*(2+.0325)) + 906*log(.25*(1-.0325)) + 904*log(.25*(1-.0325)) + 32*log(.25*(.0325))))
w
pval <- 1 - pchisq(w,1)
pval
```

c.)
```{r}
w <- -2 * ((1997*log(.25*(2+.03)) + 906*log(.25*(1-.03)) + 904*log(.25*(1-.03)) + 32*log(.25*(.03))) - (1997*log(.25*(2+.0325)) + 906*log(.25*(1-.0325)) + 904*log(.25*(1-.0325)) + 32*log(.25*(.0325))))
w
pval <- 1 - pchisq(w,1)
pval
```
d.)
```{r}
w <- -2 * ((1997*log(.25*(2+.07)) + 906*log(.25*(1-.07)) + 904*log(.25*(1-.07)) + 32*log(.25*(.07))) - (1997*log(.25*(2+.0325)) + 906*log(.25*(1-.0325)) + 904*log(.25*(1-.0325)) + 32*log(.25*(.0325))))
w
pval <- 1 - pchisq(w,1)
pval
```
explanations for b-d on paper
