---
title: "Stat 344 -- HW 7"
author: "Trey Tipton"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code

require(fastR2)

require(faraway)
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### Problem 6.6
c.)
```{r}
x <- c(1,2,3,4)
y <- c(2,5,6,8)

beta1 <- (sum(x*y))/(sum(x^2))
beta1

lm(y ~ 0 + x)
lm(y ~ -1 + x)

summary(lm(y ~ 0 + x))
```
e.)
```{r}
e <- y - (beta1*x)
sumq <- sum(e^2)
esigma <- sqrt(sumq/(length(y)-1))
esigma
se <- esigma/(sqrt(sum(x^2)))
se
```



### Problem 6.34
```{r}
model <- lm(distance ~ projectileWt, data = Trebuchet2)
x <- Trebuchet2$projectileWt
se.beta1 <- x - mean(x)
summary(model)

sse <- sum(resid(model)^2)
s <- sqrt(sse/(length(x) - 2))
s
sxx <- sum((x - mean(x))^2)
meansq <- mean(x)^2
se.beta1 <- s/sqrt(sxx)
se.beta1

se.beta0 <- s*(sqrt((1/length(x)) + (meansq/sxx)))
se.beta0

```


### Problem 6.59
a.)
Response variable: Time

Predictor variable: Concentration

I chose time as the response variable because it seems like the percentage concentrations of plasma would affect how long it takes for the blood to clot.

```{r}
gf_point(time ~ conc, data = clot)
```

b.)
```{r}
clot.model <- lm(time ~ conc, data = clot)
summary(clot.model)
```
Equation:
$\hat\beta_1$ = -0.5020
$\hat\beta_0$ = 52.5791
$\sigma$ = 21.54
$$Y_{time} = 52.5791 -0.5020x_{conc} + \epsilon$$, $\epsilon$ ~ Norm(0, 21.54)

c.)

```{r}
xbar <- mean(clot$conc)
ybar <- mean(clot$time)
sy <- sd(clot$time)
sx <- sd(clot$conc)

r <- cor(clot$time ~ clot$conc)

beta1 <- r*(sy/sx)
beta1

beta0 <- ybar - beta1*xbar
beta0

resids <- clot$time - (beta0 + beta1*clot$conc)
sum <- sum(resids^2)
sigma <- sqrt(sum/length(clot$time))
sigma
```
d.)
```{r}
ll <- function(theta, x){
  beta0 <- theta[1]
  beta1 <- theta[2]
  sigma <- theta[3]
  resids <- clot$time - (beta0 + beta1*clot$conc)
  if (sigma < 0) return(NA)
  dnorm(resids, mean = 0, sd = sigma, log = TRUE)
}


library(maxLik)
maxLik(logLik = ll, start = c(beta0 = 0, beta1 = 1, sigma = .5), x = clot)
```



  