---
title: "Stat 344 -- HW 2"
author: "Trey Tipton"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code

library(maxLik)
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### Problem 5.10
```{r}

LL <- function(theta, x){
  if (theta <0) return(NA)
  dunif(x, min = 0, max = theta, log = TRUE)
}

x <- c(1.6, 2.8, 6.2, 8.2, 8.5, 8.7)

hist(x)
```


```{r}
maxLik(LL, start = c(theta = 10), x = x)
```
The maximum likelihood estimate is $\hat{\theta} = 8.7$.

### 5.12
```{r}
mrc <- readr::read_csv('http://sldr.netlify.app/data/mrc.csv') %>%
  filter(mobility > 0)
```

```{r}
LL <- function(theta, x){
  alpha <- theta[1]
  beta <- theta[2]
  dbeta(x, shape1 = alpha, shape2 = beta, log = TRUE)
}

hist(mrc$mobility)
```

```{r, warning = FALSE}
maxLik(LL, start = c(alpha = 1, beta = 1), x = mrc$mobility)
```
Since our data ranges from zero to one with a non normal shape, it does seem that beta would be a good model to fit this data and do a maximum likelihood estimation with.


This homework took me just under an hour to finish