---
title: "Stat 344 -- HW 4"
author: "Trey Tipton"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
library(maxLik)
```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->
### Book Problem 5.18
Wald:
```{r}
odds <- 35/20
stat <- log(odds)
stat - 1.96*sqrt((stat*(1-stat))/55)
stat + 1.96*sqrt((stat*(1-stat))/55)

```
The confidence interval for the log of the odds ratio is (0.428, 0.691)

LR:
```{r}
loglik_binom <- function(theta, x, n){
  x * log(theta / (1 + theta)) + (n - x) * log(1/ (1 + theta))
}
mle_binom <- maxLik(loglik_binom, start = (odds = 1), x = 35, n = 55)
coef(mle_binom)

pb <- function(theta0){
  2 * (loglik_binom(odds, 35, 55) - loglik_binom(theta0, 35, 55)) - qchisq(0.95, df = 1)
}
lo <-uniroot(pb, c(0, odds))%>%
  value()
hi <- uniroot(pb, c(odds, 100))%>%
  value()

c(lo, hi) / (1 + c(lo, hi))
```
The 95% confidence interval is (0.447, 0.801)


### Problem 1

```{r}
gapmind <- read.csv('https://sldr.netlify.app/data/gapminder_clean.csv')%>%
  mutate(income = ifelse(income == 0, 0.0000000001, income))

Lexp <- function(theta, x){
  lambda <- theta[1]
  if (lambda < 0) return(NA)
  dexp(x, rate = lambda, log = TRUE)
}

mle <- maxLik(Lexp, start = c(lambda = 1/1000), x = gapmind$income)
mle
```

$\hat{\lambda_{MLE}} = 0.0001646$


### Problem 2
Wald:
```{r}
0.0001646 - 1.96*(0.0001646/sqrt(length(gapmind$income)))
0.0001646 + 1.96*(0.0001646/sqrt(length(gapmind$income)))
```
The 95% confidence interval is (0.0001595, 0.0001697)

LR:
```{r}
income <- gapmind$income
loglik <- function(theta, n=length(income)){
  length(income) * log(theta) - theta * sum(income)
}
p <- function(t0){
  stat <- 2*(loglik(coef(mle)) - loglik(t0))
  1 - pchisq(stat, df = 1)
}
uniroot(function(t){p(t)-0.05}, c(0, coef(mle))) %>%
  value()
uniroot(function(t){p(t)-0.05}, c(10, coef(mle))) %>%
  value()

```
The 95% confidence interval is (6.1e-05, 0.000237)


The 95% confidence interval through the likelihood ratio is much wider and is not symmetric like the Wald.


### Problem 3
Hypotheses:
$H_0$: $\Omega$ = $\Omega_0$
$H_a$: $\Omega$ = $\Omega_1$
where $\theta$ = <alpha, lambda>
$\Omega_0$ = {$\theta$ | alpha = 1}
$\Omega_1$ = {$\theta$}

Test Stat:
```{r}
ll_gamma <- function(theta, x){
  alpha <- theta[1]
  lambda <- theta[2]
  if (alpha < 0) return(NA)
  if (lambda < 0) return(NA)
  sum(log(dgamma(x, shape = alpha, rate = lambda)))
}

gamma <- maxLik(logLik = ll_gamma, start = c(alpha = 1, lambda = .0001), x = income)


exp <- maxLik(logLik = ll_gamma, start = c(alpha = 1, lambda = .0001), fixed = 1, x = income)

gamma
exp


w <- 2*(-38162.53 - -38282.38)
1 - pchisq(w, 1)
```
P-value and conclusion:

With a p-value of 0 we reject the null hypothesis that an exponential is the best fit. This means we have enough evidence to say that two paramters is a better fit and gamma is better.
