---
title: "Stat 344 - RMarkdown Challenge"
author: "Trey Tipton"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(fastR2)   # this loads mosaic, ggformula, etc. too

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code


olympics <- readr::read_csv('https://sldr.netlify.com/data/athlete_events.csv')

```


<!-- Some macros to make mathematics easier -->

\newcommand{\Prob}{\mathrm{P}}
\newcommand{\intersect}{\;\cap\;}
\newcommand{\union}{\operatorname{\cup}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

### Mathematical Notation
The probability of both event A and event B happening is $P(A \cap B)$, and if A and B are independent, this quantity is equivalent to $P(A)P(B)$.

One way of writing the equation for a simple linear regression of y as a function of x is:

$$y = \beta_0 + \beta_1 x + \epsilon,$$

$$\epsilon \sim N(0,\sigma)$$

### A Dataset

A dataset on participants in the Olympic Games is available at https://sldr.netlify.com/data/athlete_events.csv.

A glimpse() of the dataset is below:
```{r}
glimpse(olympics)
```

### Graphics
## Age distribution of Olympians
```{r echo = FALSE, warning = FALSE}
gf_dhistogram(~Age, data = olympics)
```

### Numbers of Male and Female Olympians
```{r echo = FALSE}
gf_props(~Sex, data = olympics)%>%
  gf_labs(title = 'Olympians by Sex (1896 - 2016)')
```


  
