---
title: "DATA 202 Homework 6"
author: "Trey Tipton"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
# Set some useful chunk options for code chunks.
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,    # display errors but keep going.
  comment = "",    # don't add '##' characters before output.
  message = FALSE  # don't include messages in the knitted output (check them in RStudio)
  )
library(tidyverse)
library(tidymodels)
library(parsnip)
library(recipes)
library(rsample)
library(yardstick)
library(tune)
theme_set(theme_bw())
```

## Predicting the Sale Price of Homes

### 1. Setting up the problem:

a. Some features for this task might be the location, square-footage, and year of sale of the home.

b. A target for this task would be the sale price of the home.

c. This is a regression task.

d. Mean absolute error

e. Predicted values tend to be off by about 10%.

f. An inappropriate error metric would be accuracy because this is a regression model not a classification.


### 2. Fitting and evaluating a model

```{r}
ames_home_sales <- read_builtin("ames", package = "modeldata") %>% 
  mutate(Sale_Price = Sale_Price / 1000) %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal")
```

a.
```{r}
set.seed(032601)
ames_split <- initial_split(ames_home_sales, prop = .9)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

b.
```{r}
my_model_spec <- parsnip::linear_reg(mode = "regression")
my_trained_model <- my_model_spec %>% 
  fit(Sale_Price ~ Gr_Liv_Area, data = ames_train)
```

c.
```{r}
metrics <- yardstick::metric_set(mae, mape, rsq_trad)
train_predictions <- my_trained_model %>% 
  predict(ames_test) %>% 
  bind_cols(ames_test)%>% 
  metrics(truth = Sale_Price, estimate = .pred)

train_predictions
```

d. Predictions for this model tend to be off by around 20%.


### 3. Setting up the problem again:

a. Some features for this task might be the location, square-footage, and year of sale of the home.

b. A target for this task would be the sale category of the home.

c. This is a classification task.

d. Accuracy

e. Predicted values tend to be off by about 10%.

f. An inappropriate error metric would be mae because this is a classification model not a regression.


### 4. Fitting and evaluating a model

```{r}
ames_vs_median <- ames_home_sales %>% 
  mutate(sale_category = case_when(
    Sale_Price > median(Sale_Price) ~ "Above Median",
    TRUE ~ "Below Median"
  ) %>% 
    as_factor() %>% 
    fct_relevel("Above Median") # Make sure that "Above Median" is considered the positive class.
  )
```


a.
```{r}
set.seed(032601)
ames_split2 <- initial_split(ames_vs_median, prop = .9)
ames_train2 <- training(ames_split2)
ames_test2 <- testing(ames_split2)
```

b.
```{r}
my_model_spec2 <- parsnip::decision_tree(mode = "classification")
my_trained_model2 <- my_model_spec2 %>% 
  fit(sale_category ~ Gr_Liv_Area, data = ames_train2)
```

c.
```{r}
metrics <- yardstick::metric_set(accuracy, specificity, sensitivity)

train_predictions2 <- my_trained_model2 %>% 
  predict(ames_test2) %>% 
  bind_cols(ames_test2)%>% 
  metrics(truth = sale_category, estimate = .pred_class)

train_predictions2
```

d. This model predicting whether the house is above or below the median is correct about 81% of the time.




