---
title: "DATA 202 Homework 4"
author: "Trey Tipton"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
# Set some useful chunk options for code chunks.
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,    # display errors but keep going.
  comment = "",    # don't add '##' characters before output.
  message = FALSE  # don't include messages in the knitted output (check them in RStudio)
  )
library(tidyverse)
theme_set(theme_bw())
```

## The Purpose

Our goal is to do an EDA on world populations by country.

## Loading Data

```{r}
codes <- read_csv("data/country-codes.csv")

pop <- read_csv("data/country-population.csv")
```

## Querying Raw Population Data

Create a table that plots country populations over time for subsets (of you choice) of the years and of the countries. Get the country names from the other table.

```{r}
pop2 <- pop %>% 
  pivot_longer(-country, names_to = 'Year', values_to = 'Population') %>% 
  na.omit(country)

countries <- pop2 %>% 
  left_join(codes, by = c('country' = 'code')) %>% 
  rename('Country' = 'country.y') %>% 
  mutate(Year = as.numeric(Year))


countries %>% 
  # Filter for countries that start with the letter R
  filter(country == c('rou', 'rwa', 'rus')) %>%  
  # Filter for the 20th century
  filter(Year > 1899) %>%
  filter(Year < 2000) %>% 
  ggplot() +
  aes(x = Year, y = Population, color = Country) +
  geom_line() +
  labs(title = 'Population in the 20th Century of Countries that Start with R', x = 'Year', y = 'Population', 
       color = 'Country')


```

```{r}
countries %>% 
  # Filter for countries that start with the letter R
  filter(country == c('rou', 'rwa', 'rus')) %>%  
  pivot_wider(names_from = Year, values_from = Population) %>% 
  select(Country, '1900', '1910', '1920', '1997','1998', '1999') %>% 
  knitr::kable()
  
```


For the countries that start with the letter R, Russia begins with the highest population and also has the highest increase over the 20th century. Rwanda and Romania are consistently parallel, with little change over this time period; Rwanda has the lowest population of all three.


## Exploring the Data

The World Cup has taken place every 4 years since 1930, and 8 different Countries have won in throughout those years. We're gonna take a look at those 8 teams and attempt to answer the question: Do teams with more World Cup trophies have a higher population?

```{r}
countries %>% 
  # Filter for countries that start with the letter R
  filter(Country == c('Brazil', 'Italy', 'Germany', 'Argentina', 'Uruguay', 'France', 'Spain', 'England')) %>% 
  # Filter for the around 1930 to present
  filter(Year > 1875) %>%
  ggplot() +
  aes(x = Year, y = Population, color = Country) +
  geom_line() +
  labs(title = 'Population of World Cup winning Countries over Time', x = 'Year', y = 'Population', 
       color = 'Country')
```

Brazil has the largest population by the year 2010, and they had by far the largest increase since the World Cup started in 1930. They have also won the most World Cup trophies, with 5. Germany and Italy are close behind with 4 World Cup trophies each, and for most of the 20th century, they were also 2nd and 3rd in population. So far, the answer to our question seems to be yes, that teams with more World Cup trophies have a higher population (of Countries that have won). Argentina and especially Uruguay seem to be in opposition of this answer; Uruguay has by far the lowest population of all of these countries (and Argentina has the second least population), yet they both have 2 World Cup victories each, making them 4th and 5th among the 8 world Cup winners. Overall, population seems to be a mild indicator of World Cup victories.

But let's look at the country with the most wins, Brazil, in comparison with largely populated countries that also have participated in the World Cup, but not won. 

```{r}
countries %>% 
  # Filter for countries that start with the letter R
  filter(Country == c('Brazil', 'United States', 'China', 'Australia', 'India')) %>% 
  # Filter for the around 1930 to present
  filter(Year > 1875) %>%
  ggplot() +
  aes(x = Year, y = Population, color = Country) +
  geom_line() +
  labs(title = 'Population of Brazil compared to', x = 'Year', y = 'Population', 
       color = 'Country')
```

The United States, China, and India have consistently all had a higher population than Brazil, yet none of them have ever won the World Cup (or ever come close, might I add), so there is clearly other things going on that may be better indicators of World Cup victories. Things that might affect this are culture, region, accessibility, etc. Yes, countries with higher populations may have a higher chance of having better athletes, but the over sports and athletic culture in that country can affect whether that country produces good soccer players. But for the countries that do prioritize soccer (countries that win consistently), I would still say that population is a mild indicator, but there are other factors at play.











